Epistemic Semantics in Guarded String Models

### 2020-06-22 tex/draft.tex
Edits through Section 2.

\begin{example}  \label{raise1}
Let a set of guarded strings $W$ over events P and primitive tests B, and a relation $\hat{R}$ on P be given.
The corresponding relation $R$ on $W$ holds between a guarded string
\[
\alpha_0e_0 ... e_n\alpha_{n+1} 
\]
in $W$ and a guarded string $q$ in $W$ iff $q$ is of the form

@@

where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{R}$.

@@

This requires that in the alternative world $q$, each constituent event $e'_i$ is an alternative
to the event $e_i$ in the base world according to $R$.  Compatibilities between events in the alternative
world are enforced by the general definition of worlds. Consider a scenario like the one from Figure 1, with
an additional agent Cem. The base world $T b_0 c_0$ is one where the coin is tails, and first Bob looks at tails,
and them Cem looks at tails. The first event $b_0$ has the alternatives $b_0$ and $b_1$ for Aly,
and the second event $c_0$ has the alternatives $c_0$ and $c_1$ for Aly.
This results in the combinations $b_0c_0$, $b_0c_1$, $b_1c_0$, $b_1c_1$. But these
are filtered by post- and pre-conditions of events in the alternative world, so that
the set of alernatives for Aly in $T b_0 c_0$ is  $\{T b_0 c_0,H b_1 c_1\}$,
with two world-alternatives instead of four.

\section{A Language for Epistemic Sets of Guarded Strings}

To reason about epistemic alternatives, guarded strings are used to
represent different possible worlds, so agents' knowledge/beliefs are
represented by sets of guarded strings. The standard algebra for
manipulating guarded strings is Kleene Algebra with Tests
(\textsc{KAT}), which has the algebraic signature
$\tuple{K,+,;,*,\bar{},0,1}$ \cite{kozen2001automata}. However, the
KATs don't permit native reasoning about modal operators, so Epik's
signature needs to subsume the KAT operators.

Figure~\ref{fig:epik-grammar} depicts the syntax of Epik's formal term
language. As in KAT, $0$ is the null element, $1$ is the unit element,
$p + q$ indicates disjuction, $p;q$ (or simply $pq$) indicates
sequence, the boolean operators for tests $\varphi$ are the standard
ones (as in Figure~\ref{fig:effect-syntax}), and ${p}^*$ indicates
iteration of $p$. We add some additional operations to reason about epistemic
alternatives: the term $\Diamond_a~p$ represents epistemic possibility
for agent $a$ in worlds $p$, and complement $\neg p$ represents the
worlds not described by $p$. We can encode $\Box_a~p$, i.e. epistemic
certainty for agent $a$ in worlds $p$, by $\Box_a~p =
(\Diamond_a~p^c)^c$.

@@

@@

>>>>

\begin{example}  \label{raise1}
Let a set of guarded strings $W$ over events P and primitive tests B, and a relation $\hat{R}$ on P be given.
The corresponding relation $R$ holds between a guarded string
\[
\alpha_0e_0 ... e_n\alpha_{n+1} 
\]
in $W$ and a guarded string  $q$ iff $q$ is an element of $W$ and is of the form

@@

where for $0 \leq n$, $\element{\tuple{e_i,e'_i}}{\hat{R}}$.

@@

This requires that in the alternative world $q$, each constituent event $e'_i$ is an
alternative to the event $e_i$ in the base world according to $\hat{R}$.
Compatibilities between events in the alternative world are enforced by the
requirement that the alternative $q$ is an element of $W$, so that 
state and effect formulas are enforced. Consider a scenario like the one from
Figure 1, with an additional agent Cem. The base world $T b_0 c_0$ is one where the
coin is tails, and first Bob looks at tails, and them Cem looks at tails. The first
event $b_0$ has the alternatives $b_0$ and $b_1$ for Aly, and the second event $c_0$
has the alternatives $c_0$ and $c_1$ for Aly.  This results in the combinations
$b_0c_0$, $b_0c_1$, $b_1c_0$, and $b_1c_1$. But these are filtered by post- and
pre-conditions of events in the alternative world, so that the set of alernatives for
Aly in $T b_0 c_0$ is $\{T b_0 c_0,H b_1 c_1\}$, with two world-alternatives instead
of four.

\section{The language of Epistemic KAT}
In addition to a set of possible worlds and an algebra of propositions, the
application in natural language requires a logical language for naming propositions.
The standard language for Kleene algebra with tests has the signature
$\tuple{K,+,\cdot,*,\bar{},0,1}$ \citep{kozen2001automata}.  In a guarded string
model for KAT, $K$ is a set of guarded strings, $+$ is set union, $\cdot$ is fusion
product raised to sets, $*$ is Kleene star, $\bar{}$ is complement for tests, $0$ the
empty set (functioning as an identity for $0$, and as False for tests) and 1 is the
set of atoms (functioning as the identity element for $\cdot$, and as True for
tests).  To this signature we add a unary modal operation $\Diamond_a$ for each
agent, and a unary complement operation $\not$ on elements of $K$. The latter is
included because natural languages have sentence negation, and it should be possible
to represent the propositional complement of a formula that captures the semantics of
a natural language sentence.  In addition, universal box modalities are defined as
duals of existential diamond modalities.

With modalities and propositional negation added, the signature of $n$-agent
epistemic KAT is $\tuple{K,+,\cdot,*,\bar{},0,1,\neg,\Diamond_1
  ... \Diamond_n}$. Figure \ref{termfigure} defines the syntax of the language.
\ex{interpretation1} illustrates the application of the language with some natural
language sentences, and their representations in the logical language.  While there
is no constant for the set of worlds, the term $\neg 0$ (the complement of the empty
set) can be used as one.  Where $h$ is the primitive test ``it' heads'' from the coin
example, $(\neg 0)h$ designates the set of worlds where it is heads.  The reason is
that $h$ corresponds to the set of atoms that evaluate $h$ to true. Then the product
term $(\neg 0)h$ restricts the set of worlds to the set of guarded strings that end
with and atom that evaluate $h$ to True.

\begin{example}   \label{interpretation1}
\begin{tabular}[t]{ll}
  It's heads. & It's tails. \\
  $(\neg 0)h$ & $(\neg 0)t$ \\ \hline
\end{tabular}
\end{example}  
\vspace{-3mm}
\hspace{2mm}\begin{tabular}{l}
Amy doesn't know that it's heads.  \\
$\neg {\Box}_a (\neg 0)h$\\ \hline
Amy knows whether it's tails. \\
$\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t$ \\ \hline
It's heads and Bob knows that it's tails. \\
$(\neg 0)h + \Box_a(\neg 0)\neg t$ \\  \hline
Bob knows that Amy knows whether it's tails. \\
$\Box_b(\Box_a(\neg 0)t + \Box_a(\neg 0)\neg t)$ \\ \hline
\end{tabular}
%% (\neg 0)t

@@

  {\em Consider adding the syntax of tests including $\bar{}$. Consider making product implicit,
    in place of $\mathbin{;}$.}

@@

{\em This stuff was at the end of Section 2, it became redundant. Also it refers to the interpretation function, which is introduced in Section 3.}
